---
title: "Regular Expressions Lab"
author: "Scott Karr"
date: "February 14, 2016"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---
######--------------------------------------------------------------
##### AUTOMATED DATA COLLECTION WITH R
######  SIMON MUNZERT, CHRISTIAN RUBBA, PETER MEISSNER, DOMINIC NYHUIS
######  CODE CHAPTER 8: REGULAR EXPRESSIONS AND ESSENTIAL STRING FUNCTIONS
######  Problem 3,7,8,9
######--------------------------------------------------------------
(a) Load Packages
```{r load-pkgs, eval=TRUE}
library(stringr)
library(XML)
library(RCurl)
library(bitops)
library(tau)
```

(b) A Difficult Example
```{r example, eval=TRUE}
##start by assigning a raw data vector of one element which mixes phone and name data
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5543642Dr. Julius Hibbert"
```

```{r extract-information, eval=TRUE}
##extract a vector of name elements 
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
##extract a vector of phone elements
phone <- unlist(str_extract_all(raw.data, "\\(?(\\d{3})?\\)?(-| )?\\d{3}(-| )?\\d{4}"))
phone
##assign elements to dataframe
data.frame(name = name, phone = phone)
```

(c) Building REGEX

####8.1 Regular expressions
```{r a-running-example, eval=TRUE}
##assign a single value string vector
example.obj <- "1. A small sentence. - 2. Another tiny sentence."
```

#####8.1.1 Exact character matching
```{r a-string_matching-itself, eval=TRUE}
##extract string matching itself with 1st match
str_extract(example.obj, "small")
##extract string without match
str_extract(example.obj, "banana")
##extract string matching itself with all matches
unlist(str_extract_all(example.obj, "sentence"))
```

```{r string-vector, eval=TRUE}
##extract a string vector with 3 elements
out <- str_extract_all(c("text", "manipulation", "basics"), "a")
##display
out
```

```{r case-sensitive-matching, eval=TRUE}
##extract match on 1st small
str_extract(example.obj, "small")
##extract string without match on 1st case
str_extract(example.obj, "SMALL")
##extract string override match on 1st case
str_extract(example.obj, ignore.case("SMALL"))
```

```{r match-a-particle, eval=TRUE}
##extract matching a particle, alphabetic characters and blank spaces
unlist(str_extract_all(example.obj, "en"))
##extract string on 1st match on 1st case
str_extract(example.obj, "mall sent")
```

```{r match-beginning-end, eval=TRUE}
##extract match on 1st case
str_extract(example.obj, "2")
##extract match on beginning case
str_extract(example.obj, "^2")
##extract match on end case
unlist(str_extract_all(example.obj, "sentence$"))
```

```{r pipe-operator, eval=TRUE}
##use of pipe operator as "or"
unlist(str_extract_all(example.obj, "tiny|sentence"))
```

#####8.1.2 Generalizing regular expressions  
```{r any-character, eval=TRUE}
##extract using any character regex "sm.ll"
str_extract(example.obj, "sm.ll")
```

```{r character-classes, eval=TRUE}
##extract using character-class regex "sm[abc]ll"
str_extract(example.obj, "sm[abc]ll")
```

```{r character-class-range, eval=TRUE}
##extract using character-class-range regex "sm[a-p]ll"
str_extract(example.obj, "sm[a-p]ll")
```

```{r literal, eval=TRUE}
##extract using literal dots inside a character class regex "[uvw. ]"
unlist(str_extract_all(example.obj, "[uvw. ]"))
```

```{r predefined-char-class, eval=TRUE}
##extract using predefined character classes
unlist(str_extract_all(example.obj, "[[:punct:]]"))
```

```{r predefined-char-class2, eval=TRUE}
##wrong usage of predefined character classes
unlist(str_extract_all(example.obj, "[:punct:]"))
##      redundancy inside a character class
unlist(str_extract_all(example.obj, "[AAAAA]"))
##special characters inside a character class range
str_extract("François Hollande", "Fran[a-z]ois")
## w/ alpha char matching
str_extract("François Hollande", "Fran[[:alpha:]]ois")
## w/ punct char matching
unlist(str_extract_all(example.obj, "[[:punct:]ABC]"))
##negation of a character class
unlist(str_extract_all(example.obj, "[^[:alnum:]]"))
```

```{r quantifiers, eval=TRUE}
##Quantifiers
str_extract(example.obj, "s[[:alpha:]][[:alpha:]][[:alpha:]]l")
str_extract(example.obj, "s[[:alpha:]]{3}l")
```

```{r greedy-matching, eval=TRUE}
##Greedy matching
str_extract(example.obj, "A.+sentence")
str_extract(example.obj, "A.+?sentence")
```

```{r quantifiers2, eval=TRUE}
##Quantifying a group of characters
unlist(str_extract_all(example.obj, "(.en){1,5}"))
unlist(str_extract_all(example.obj, ".en{1,5}"))
```

```{r metacharacters, eval=TRUE}
##Literal metacharacters
unlist(str_extract_all(example.obj, "\\."))
unlist(str_extract_all(example.obj, fixed(".")))
```

```{r metacharacters2, eval=TRUE}
##Shortcuts of character classes
unlist(str_extract_all(example.obj, "\\w+"))
```

```{r metacharacters3, eval=TRUE}
##Word edges
unlist(str_extract_all(example.obj, "e\\>"))
unlist(str_extract_all(example.obj, "e\\b"))
```

```{r metacharacters4, eval=TRUE}
##Backreferencing
str_extract(example.obj, "([[:alpha:]]).+?\\1")
str_extract(example.obj, "(\\<[b-z]+\\>).+?\\1")
```

#####8.1.3 The introductory example reconsidered  
```{r example2, eval=TRUE}
##assign data to vector
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5543642Dr. Julius Hibbert"
##extract names with "[[:alpha:]., ]{2,}"
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
##display
name
##extract phones with "\\(?(\\d{3})?\\)?(-| )?\\d{3}(-| )?\\d{4}"
phone <- unlist(str_extract_all(raw.data, "\\(?(\\d{3})?\\)?(-| )?\\d{3}(-| )?\\d{4}"))
##display
phone
```

(d) Stringr Package

####8.2 String Processing
#####8.2.1 The stringr package
```{r stringr, eval=TRUE}
##str_extract()
str_extract(example.obj, "tiny")
str_extract_all(example.obj, "[[:digit:]]")
```

```{r stringr2, eval=TRUE}
##str_locate() 
str_locate(example.obj, "tiny")
```

```{r stringr3, eval=TRUE}
##str_sub()
str_sub(example.obj, start = 35, end = 38)
str_sub(example.obj, 35, 38) <- "huge"
example.obj
```

```{r stringr4, eval=TRUE}
##str_replace()
str_replace(example.obj, pattern = "huge", replacement = "giant")
```

```{r stringr5, eval=TRUE}
##str_split() / str_split_fixed()
unlist(str_split(example.obj, "-"))
as.character(str_split_fixed(example.obj, "[[:blank:]]", 5))
```

```{r stringr6, eval=TRUE}
##A character vector
char.vec <- c("this", "and this", "and that")
```

```{r stringr7, eval=TRUE}
##str_detect()
str_detect(char.vec, "this")
```

```{r stringr8, eval=TRUE}
##str_count()
str_count(char.vec, "this")
str_count(char.vec, "\\w+")
```

```{r stringr9, eval=TRUE}
##str_dup()
dup.obj <- str_dup(char.vec, 3)
dup.obj
```

```{r stringr10, eval=TRUE}
##str_length()
length.char.vec <- str_length(char.vec)
length.char.vec
```

```{r stringr11, eval=TRUE}
##str_pad() / str_trim()
char.vec <- str_pad(char.vec, width = max(length.char.vec), side = "both", pad = " ")
char.vec
char.vec <- str_trim(char.vec)
char.vec
```

```{r stringr12, eval=TRUE}
##str_c()
cat(str_c(char.vec, collapse = "\n"))
str_c("text", "manipulation", sep = " ")
str_c("text", c("manipulation", "basics"), sep = " ")
```

```{r functions, eval=TRUE}
##8.2.2 A couple more handy functions
##Approximate matching
agrep("Barack Obama", "Barack H. Obama", max.distance = list(all = 3))
agrep("Barack Obama", "Michelle Obama", max.distance = list(all = 3))
```

```{r functions2, eval=TRUE}
##Partial string matching
pmatch(c("and this", "and that", "and these", "and those"), char.vec)
```

```{r functions3, eval=TRUE}
##Make unique
make.unique(c("a", "b", "a", "c", "b", "a"))
```


load("episodes.Rdata")
grep("Homer", episodes$title[1:10], value = T)
grepl("Homer", episodes$title[1:10])
iffer1 <- grepl("Homer", episodes$title)
iffer2 <- grepl("Lisa", episodes$title)
iffer <- iffer1 & iffer2
episodes$title[iffer]

grepall <- function(pattern, x, ignore.case = FALSE, perl = FALSE, fixed = FALSE, useBytes = FALSE, value = FALSE, logic = FALSE){
    error and exception handling
    if(length(pattern) == 0 | length(x) == 0){
        warning("Length of pattern or data equals zero.")
        return(NULL)
    }
    apply grepl() and all()
    indices <- sapply(pattern, grepl, x, ignore.case, perl, fixed, useBytes)
    index <- apply(indices, 1, all)
    indexation and return of results
    if(logic == T) return(index)
    if(value == F) return((1:length(x))[index])
    if(value == T) return(x[index])
}
grepall(c("Lisa", "Homer"), episodes$title)
grepall(c("Lisa", "Homer"), episodes$title, value = T)
```

(e) Additional analysis

####8.3 A word on character encodings  
Query your locale
Sys.getlocale()

An example string
small.frogs <- "Små grodorna, små grodorna är lustiga att se."
small.frogs

Translate the encoding
small.frogs.utf8 <- iconv(small.frogs, from = "windows-1252", to = "UTF-8")
small.frogs.utf8

Declare an encoding
Encoding(small.frogs.utf8) <- "windows-1252"
small.frogs.utf8

Sample from the list of available conversions
sample(iconvlist(), 10)

Meta tags from Science
enc.test <- getURL("http://www.sciencemag.org/")
unlist(str_extract_all(enc.test, "<meta.+?>"))

Guess encodings
is.locale(small.frogs)
is.ascii(small.frogs)
