---
title: "Regular Expressions Homework"
author: "Scott Karr"
date: "February 28, 2016"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---
######--------------------------------------------------------------
##### AUTOMATED DATA COLLECTION WITH R
######  CODE CHAPTER 8: REGULAR EXPRESSIONS AND ESSENTIAL STRING FUNCTIONS
######  Problems 3,7,8,9
######--------------------------------------------------------------
(a) Load Packages
```{r load-pkgs, eval=TRUE}
library(stringr)
library(XML)
library(RCurl)
library(bitops)
library(tau)
```

(b) Problem #3  
Copy the introductory example.  The vector name stroes the extracted names.
```{r example, eval=TRUE}
##start by assigning a raw data vector of one element which mixes phone and name data
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5543642Dr. Julius Hibbert"
```

```{r extract-information, eval=TRUE}
##extract a vector of name elements 
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
name
##extract a vector of phone elements
phone <- unlist(str_extract_all(raw.data, "\\(?(\\d{3})?\\)?(-| )?\\d{3}(-| )?\\d{4}"))
phone <- unlist(str_extract_all(raw.data, "\\(?(\\d{3})?\\)?(-| )?\\d{3}(-| )?\\d{4}"))
phone
##assign elements to dataframe
data.frame(name = name, phone = phone)
```

#####     (a) Use the tools of the chapter to rearrange the vector so that all elements conform to the standard first_name last_name.
```{r reformat-names, eval=TRUE}
##extract a vector of fname elements 
fname <- unlist(str_extract_all(name, "[:punct:] [[:alpha:]]{2,}$|[[:alpha:]]{2,} "))
fname
fname <- unlist(str_extract_all(fname,"[[:alpha:]]{2,}"))
fname
##extract a vector of lname elements 
lname <- unlist(str_extract_all(name, "[^[:punct:]] [[:alpha:]]{2,}$|[[:alpha:]]{2,}, "))
lname <- unlist(str_extract_all(lname,"[[:alpha:]]{2,}"))
lname
##extract a vector of title elements 
title <- unlist(str_extract_all(name, "[[:alpha:]]{2,}\\."))
title
title_exists <- unlist(str_detect(name, title))
title_exists
##update the name dataframe
data.frame(fname=fname, lname=lname, title_exists=title_exists)
fname2 <- unlist(str_extract_all(name, " [[:alpha:]]{1}\\.? [[:alpha:]]{1,}\\.?"))
fname2 <- unlist(str_extract_all(fname2,"[[:alpha:]]{1}\\.? [[:alpha:]]{1,}\\.?"))
fname2
fname2_exists <- unlist(str_detect(name, " [[:alpha:]]{1}\\.? [[:alpha:]]{1,}\\.?"))
fname2_exists
data.frame(fname=fname, lname=lname, phone=phone, title_exists=title_exists, fname2_exists=fname2_exists)
```
(c) Problem #7
The expression "<.+>" doesn't repeat the alpha characters contained in the tag 
```{r html,eval=TRUE}
raw.data <- "<title>+++BREAKING NEWS+++</title>"
html <- str_extract(raw.data,"<[[:alpha:]]{1,}>")
html
```

(d) Problem #8
The expression [^1-9=+*()] is replaced with "[^[1-9]{1,}]" which captures =+*()
```{r binomial,eval=TRUE}
raw.data <- "(5-3)^2=5^2-2*5*3+3^2"
binomial <- unlist(str_extract_all(raw.data, "[^[1-9]{1,}]"))
binomial
```


